---
title: "Social Web Analytics Projet"
author: "Phuong Linh Le"
date: "2026-10-05"
output:
  pdf_document:
    latex_engine: "xelatex"
  html_document:
    df_print: paged
---

```{r, eval=FALSE}
# Load the package used for Bluesky data collection
library(atrrr)
```

```{r, eval=FALSE}
# Initial API test - Retrieve a small sample to inspect the Bluesky data structure
alfred_test <- search_skeet(
  "Cyclone Alfred",
  sort = "latest",
  since = "2025-02-23",
  limit = 100
)

# Inspect the dimensions and available variables in the initial sample
dim(alfred_test)
names(alfred_test)

# Inspect key variables relevant to the project
head(alfred_test[, c(
  "created_at",
  "author_handle",
  "text",
  "reply_count",
  "repost_count",
  "like_count"
)])

# Check the actual temporal coverage returned by the API
range(alfred_test$created_at)
head(sort(alfred_test$created_at))
tail(sort(alfred_test$created_at))
```

```{r, eval=FALSE}
# Inspect the available arguments of search_skeet before defining the collection period
args(search_skeet)
```

```{r, eval=FALSE}
# Test retrieval across the full event period using a single API request
alfred_event <- search_skeet(
  "Cyclone Alfred",
  sort = "latest",
  since = "2025-02-23",
  until = "2025-03-17",
  limit = 500
)

# Examine the coverage of the initial event-level retrieval
dim(alfred_event)
range(alfred_event$created_at, na.rm = TRUE)
length(unique(alfred_event$author_handle))
table(as.Date(alfred_event$created_at))
```

```{r, eval=FALSE}
# 1. Check the dataset size
dim(alfred_event)
# 2. Check the actual date range
range(alfred_event$created_at, na.rm = TRUE)
# 3. Check the number of unique authors
length(unique(alfred_event$author_handle))
# 4. Check the number of posts per day
table(as.Date(alfred_event$created_at))
```

```{r, eval=FALSE}
# Check the number of posts that are replies
sum(!is.na(alfred_event$in_reply_to))
# Calculate the total number of replies received
sum(alfred_event$reply_count, na.rm = TRUE)
# Calculate the total number of reposts
sum(alfred_event$repost_count, na.rm = TRUE)
# Calculate the total number of quotes
sum(alfred_event$quote_count, na.rm = TRUE)
# Check the number of posts containing mentions
sum(lengths(alfred_event$mentions) > 0, na.rm = TRUE)
# Check the number of posts containing links
sum(lengths(alfred_event$links) > 0, na.rm = TRUE)
```

```{r, eval=FALSE}
# Test data retrieval for the earlier period of the event
alfred_before <- search_skeet(
  "Cyclone Alfred",
  sort = "latest",
  since = "2025-02-23",
  until = "2025-03-05",
  limit = 500
)
```
```{r, eval=FALSE}
# Check the dataset size
dim(alfred_before)
# Check the actual date range returned
range(alfred_before$created_at, na.rm = TRUE)
# Check the number of unique authors
length(unique(alfred_before$author_handle))
```
```{r, eval=FALSE}
# Test data retrieval using a one-day search window
alfred_23feb <- search_skeet(
  "Cyclone Alfred",
  sort = "latest",
  since = "2025-02-23",
  until = "2025-02-24",
  limit = 500
)
```
```{r, eval=FALSE}
# Check the dataset size
dim(alfred_23feb)
# Check the actual date range returned
range(alfred_23feb$created_at, na.rm = TRUE)
# Check the number of unique authors
length(unique(alfred_23feb$author_handle))
```

```{r, eval=FALSE}
# Create a sequence of dates covering the full study period
dates <- seq(
  as.Date("2025-02-23"),
  as.Date("2025-03-16"),
  by = "day"
)
# Create an empty list to store the daily search results
alfred_daily <- vector("list", length(dates))
# Collect Bluesky posts separately for each day
for (i in seq_along(dates)) {
  # Define the start and end date for each daily search
  start_date <- dates[i]
  end_date <- start_date + 1
  # Display the date currently being collected
  cat("Collecting:", as.character(start_date), "\n")
  # Retrieve posts containing the search term "Cyclone Alfred"
  alfred_daily[[i]] <- search_skeet(
    "Cyclone Alfred",
    sort = "latest",
    since = as.character(start_date),
    until = as.character(end_date),
    limit = 1000
  )
}
```

```{r, eval=FALSE}
# Load dplyr for data manipulation
library(dplyr)
# Combine all daily search results into one final dataset
alfred_all <- bind_rows(alfred_daily)
# Check the size of the final dataset
dim(alfred_all)
# Check the actual date range of the final dataset
range(alfred_all$created_at, na.rm = TRUE)
# Check the number of unique authors
length(unique(alfred_all$author_handle))
# Check for duplicate posts based on URI
sum(duplicated(alfred_all$uri))
```

```{r, eval=FALSE}
# Check the number of posts collected per day
table(as.Date(alfred_all$created_at))
```

```{r, eval=FALSE}
# Check the number of posts that are replies
sum(!is.na(alfred_all$in_reply_to))
# Calculate the total number of replies received
sum(alfred_all$reply_count, na.rm = TRUE)
# Calculate the total number of reposts
sum(alfred_all$repost_count, na.rm = TRUE)
# Calculate the total number of quotes
sum(alfred_all$quote_count, na.rm = TRUE)
# Check the number of posts containing mentions
sum(lengths(alfred_all$mentions) > 0, na.rm = TRUE)
# Check the number of posts containing links
sum(lengths(alfred_all$links) > 0, na.rm = TRUE)
```

```{r, eval=FALSE}
# Save the final collected dataset for subsequent analysis
file.exists("alfred_all.rds")
```
# Data Collection
# Data Understanding and Preparation
## Step 1 - Load Dataset
```{r}
# Load required package
library(dplyr)
# Load the shared dataset
alfred_all_raw <- readRDS("Data/alfred_all.rds")
# Create a working copy
alfred_all <- alfred_all_raw
# Initial check
dim(alfred_all)
```
##Step 2 - Dataset Audit
```{r}
# Dataset class
class(alfred_all)
# Dataset dimensions
dim(alfred_all)
# Variable names
names(alfred_all)
# Variable types and structure
dplyr::glimpse(alfred_all)
# Inspect first few records
head(alfred_all)
# Check duplicate post identifiers
sum(duplicated(alfred_all$uri))
```
## Step 3 - Missingness and Record Suitability
```{r}
# Number of missing values in each variable
missing_values <- colSums(is.na(alfred_all))
missing_values
sum(is.na(alfred_all$uri))
sum(is.na(alfred_all$author_handle))
sum(is.na(alfred_all$text))
sum(is.na(alfred_all$created_at))
sum(is.na(alfred_all$in_reply_to))
sum(is.na(alfred_all$quotes))
```
## Step 4 - User and Temporal Coverage
```{r}
# Number of unique authors
unique_users <- length(unique(alfred_all$author_handle))
unique_users
# Check created_at type
class(alfred_all$created_at)
# Actual post creation time range
date_range <- range(
  alfred_all$created_at,
  na.rm = TRUE
)
date_range
# Number of posts by date
posts_per_day <- table(
  as.Date(alfred_all$created_at)
)
posts_per_day
```
## Step 5 - Text Quality
```{r}
# Missing text
missing_text <- sum(is.na(alfred_all$text))
missing_text
# Blank / empty text
blank_text <- sum(
  trimws(alfred_all$text) == "",
  na.rm = TRUE
)
blank_text
# Posts containing usable non-empty text
usable_text <- sum(
  !is.na(alfred_all$text) &
    trimws(alfred_all$text) != ""
)
usable_text
# Inspect language variable
class(alfred_all$langs)
str(alfred_all$langs)
# Number of language values attached to each post
table(lengths(alfred_all$langs))
# Inspect actual language codes
language_values <- unlist(
  alfred_all$langs,
  use.names = FALSE
)
sort(
  table(language_values),
  decreasing = TRUE
)
alfred_all$text_usable <-
  !is.na(alfred_all$text) &
  trimws(alfred_all$text) != ""
table(alfred_all$text_usable)
```
## Step 6 - Link Preparation
```{r}
# Inspect links variable
class(alfred_all$links)
str(alfred_all$links)
# Number of links recorded per post
table(lengths(alfred_all$links))
# Number of posts with at least one recorded link
posts_with_links <- sum(
  lengths(alfred_all$links) > 0
)
posts_with_links
link_rows <- which(
  lengths(alfred_all$links) > 0
)
head(
  alfred_all$links[link_rows],
  10
)
alfred_all$has_link <-
  lengths(alfred_all$links) > 0
table(alfred_all$has_link)
prop.table(
  table(alfred_all$has_link)
)
```
## Step 7 - Network Feasibility
### 7A - Replies
```{r}
# Number of reply posts
reply_posts <- sum(
  !is.na(alfred_all$in_reply_to)
)
reply_posts
# Inspect reply identifiers
class(alfred_all$in_reply_to)
head(
  na.omit(alfred_all$in_reply_to),
  10
)
reply_rows <- which(
  !is.na(alfred_all$in_reply_to)
)
reply_parent_match <- match(
  alfred_all$in_reply_to[reply_rows],
  alfred_all$uri
)
reply_parents_in_dataset <- sum(
  !is.na(reply_parent_match)
)
reply_parents_in_dataset
reply_mapping_rate <-
  reply_parents_in_dataset /
  length(reply_rows)
reply_mapping_rate
reply_example <- data.frame(
  source_user =
    alfred_all$author_handle[reply_rows],
  target_uri =
    alfred_all$in_reply_to[reply_rows],
  target_user =
    alfred_all$author_handle[reply_parent_match]
)
head(reply_example, 10)
```
### 7B - Mentions
```{r}
# Number of posts containing mentions
mention_posts <- sum(
  lengths(alfred_all$mentions) > 0
)
mention_posts
# Inspect structure
class(alfred_all$mentions)
str(alfred_all$mentions)
# Inspect non-empty mention examples
mention_rows <- which(
  lengths(alfred_all$mentions) > 0
)
head(
  alfred_all$mentions[mention_rows],
  10
)
head(
  unlist(
    alfred_all$mentions[mention_rows],
    use.names = FALSE
  ),
  20
)
```
### 7C - Quotes
```{r}
# Inspect quotes variable
class(alfred_all$quotes)
str(alfred_all$quotes)
# Number of posts with quote information
quote_posts <- sum(
  !is.na(alfred_all$quotes)
)
quote_posts
# Inspect examples
head(
  na.omit(alfred_all$quotes),
  10
)
if (is.character(alfred_all$quotes)) {
  quote_rows <- which(
    !is.na(alfred_all$quotes)
  )
  quote_posts_in_dataset <- sum(
    alfred_all$quotes[quote_rows] %in%
      alfred_all$uri
  )
  quote_posts_in_dataset
}
```
## Step 8 - Final Preparation Summary
```{r}
prep_summary <- data.frame(
  Metric = c(
    "Total posts",
    "Total variables",
    "Unique users",
    "Duplicate URI",
    "Missing created_at",
    "Missing text",
    "Blank text",
    "Usable text posts",
    "Posts with links",
    "Reply posts",
    "Reply parents found in dataset",
    "Posts with mentions",
    "Posts with quote information"
  ),
  Value = c(
    nrow(alfred_all),
    ncol(alfred_all),
    unique_users,
    sum(duplicated(alfred_all$uri)),
    sum(is.na(alfred_all$created_at)),
    missing_text,
    blank_text,
    usable_text,
    posts_with_links,
    reply_posts,
    reply_parents_in_dataset,
    mention_posts,
    quote_posts
  )
)
prep_summary

date_range
```

# RQ1 - Content Themes

**Research Question:** What information themes characterised Bluesky discussion during Tropical Cyclone Alfred?

## 1 - Select Usable Text

Only posts containing usable, non-empty text were selected for the content analysis.

```{r}
# Select usable text for content analysis
rq1_text <- alfred_all$text[alfred_all$text_usable]
```

## 2 - Text Preprocessing

The selected posts were converted into a text corpus and cleaned before term-frequency analysis. Unsupported symbols, URLs, numbers, punctuation, common English stopwords, and query-specific terms were removed.

```{r}
# Load the package for text processing
library(tm)
# Create a text corpus
rq1_corpus <- Corpus(VectorSource(rq1_text))
# Convert text to ASCII to remove unsupported symbols and emojis
rq1_corpus <- tm_map(
  rq1_corpus,
  content_transformer(function(x) iconv(x, to = "ASCII", sub = ""))
)
```
```{r}
# 2.1 - Remove URLs from the corpus
rq1_corpus <- tm_map(
  rq1_corpus,
  content_transformer(function(x) gsub("http\\S+|www\\S+", "", x))
)
```
```{r}
# 2.2 - Convert all text to lowercase
rq1_corpus <- tm_map(
  rq1_corpus,
  content_transformer(tolower)
)
```
```{r}
# 2.3 - Remove numbers from the corpus
rq1_corpus <- tm_map(
  rq1_corpus,
  removeNumbers
)
```
```{r}
# 2.4 - Remove punctuation from the corpus
rq1_corpus <- tm_map(
  rq1_corpus,
  removePunctuation
)
```
```{r}
# 2.5 - Remove common English stopwords
rq1_corpus <- tm_map(
  rq1_corpus,
  removeWords,
  stopwords("english")
)
```
```{r}
# 2.6 - Remove query-specific words
custom_stopwords <- c("cyclone", "alfred")

rq1_corpus <- tm_map(
  rq1_corpus,
  removeWords,
  custom_stopwords
)
```
```{r}
# 2.7 - Remove extra whitespace
rq1_corpus <- tm_map(
  rq1_corpus,
  stripWhitespace
)
```
## 3 - Initial Term-Frequency Analysis

An initial Term-Document Matrix was created to identify the most frequent terms and determine whether additional text preprocessing was required.

```{r}
# Create the initial Term-Document Matrix
rq1_tdm <- TermDocumentMatrix(rq1_corpus)
# Calculate initial term frequencies
rq1_term_frequency <- rowSums(as.matrix(rq1_tdm))
# Display the 20 most frequent terms
sort(rq1_term_frequency, decreasing = TRUE)[1:20]
```
The initial term-frequency results were used to identify potentially uninformative frequent terms and to assess whether related word forms were being counted separately. These diagnostic checks informed the final preprocessing decisions before the main results were interpreted.

### 3.1 - Inspect Potentially Uninformative Terms

Frequently occurring terms that appeared potentially uninformative were examined in their original post context before deciding whether they should be removed.

```{r}
# Inspect the context of potentially uninformative frequent terms
terms_to_check <- c("will", "now", "make", "news", "live")

for (term in terms_to_check) {
  
  # Find posts containing the selected term
  matched_posts <- rq1_text[
    grepl(paste0("\\b", term, "\\b"), tolower(rq1_text))
  ]
  
  # Display the term and three example posts
  cat("\n\n--- TERM:", term, "---\n")
  print(head(matched_posts, 3))
}
```

The contextual inspection showed that *will*, *now*, and *make* were largely generic and provided limited information about the cyclone discussion. In contrast, *news* and *live* were retained because they frequently appeared in posts sharing news coverage and real-time event updates. Therefore, only *will*, *now*, and *make* were removed as additional stopwords.

```{r}
# Remove additional uninformative words identified through context inspection
additional_stopwords <- c("will", "now", "make")

rq1_corpus <- tm_map(
  rq1_corpus,
  removeWords,
  additional_stopwords
)

# Remove extra whitespace created after removing additional words
rq1_corpus <- tm_map(
  rq1_corpus,
  stripWhitespace
)
```

### 3.2 - Check Related Word Forms

Related forms of several prominent words were examined to determine whether variations of the same concept were being counted separately.

```{r}
# Check whether related word forms are splitting term frequencies
stem_check_terms <- c(
  "flood", "flooded", "flooding",
  "warn", "warned", "warning", "warnings",
  "prepare", "prepared", "preparing",
  "impact", "impacted", "impacts"
)

rq1_term_frequency[
  intersect(stem_check_terms, names(rq1_term_frequency))
]
```
The results showed that related forms of the same words were being counted separately. For example, *flood*, *flooded*, and *flooding*, as well as different forms of *warning*, *prepare*, and *impact*, appeared as separate terms. Stemming was therefore applied to consolidate related word forms and reduce fragmentation in the term-frequency analysis.

```{r}
# Stem related word forms to a common root
rq1_corpus <- tm_map(
  rq1_corpus,
  stemDocument
)
```
```{r}

# Remove residual query terms after stemming
rq1_corpus <- tm_map(
  rq1_corpus,
  removeWords,
  c("cyclonealfr", "tcalfr", "cyclon")
)

rq1_corpus <- tm_map(rq1_corpus, stripWhitespace)

```

## 4 - Final Term-Frequency Analysis

After the final preprocessing steps, the Term-Document Matrix and term frequencies were recalculated.

```{r}
# Recreate the Term-Document Matrix after final text preprocessing
rq1_tdm <- TermDocumentMatrix(rq1_corpus)
# Recalculate term frequencies after stemming
rq1_term_frequency <- rowSums(as.matrix(rq1_tdm))
# Display the 20 most frequent terms
sort(rq1_term_frequency, decreasing = TRUE)[1:20]
```
The final term-frequency results show that Bluesky discussion was strongly focused on the cyclone's geographic location, development, and impacts. Geographic terms such as *Queensland*, *Brisbane*, *coast*, and *NSW* were among the most frequent, indicating substantial attention to areas affected or threatened by the cyclone. Weather- and event-related terms such as *tropical*, *storm*, *wind*, *landfall*, and *rain* were also prominent, reflecting continued discussion of the cyclone's movement and conditions. Terms including *flood*, *power*, and *prepare* point to discussion of potential or experienced impacts and preparedness, while *news*, *live*, and *update* suggest that real-time information sharing was also prominent.

## 5 - Visualisation

### 5.1 - Prepare Top 20 Terms

The 20 most frequent terms from the final term-frequency analysis were prepared for visualisation.

```{r}
# Prepare the Top 20 frequent terms for visualisation
rq1_top20 <- sort(
  rq1_term_frequency,
  decreasing = TRUE
)[1:20]

# Convert the Top 20 terms into a data frame
rq1_top20_df <- data.frame(
  term = names(rq1_top20),
  frequency = as.numeric(rq1_top20)
)

# Display the prepared data
rq1_top20_df
```

### 5.2 - Create Readable Labels

Because stemming produces shortened word roots, readable labels were used in the visualisations while retaining the original calculated frequencies.

```{r}
# Create readable labels for stemmed terms in visualisations
rq1_top20_df$display_term <- c(
  "Tropical",
  "Queensland",
  "Coast",
  "Brisbane",
  "Australia",
  "News",
  "Flood",
  "March",
  "NSW",
  "Live",
  "Wind",
  "Storm",
  "Landfall",
  "Ex-tropical",
  "Rain",
  "Power",
  "Update",
  "Hit",
  "South",
  "Prepare"
)

# Check that display labels match the original stemmed terms
rq1_top20_df[, c("term", "display_term", "frequency")]
```

### 5.3 - Top 20 Term-Frequency Bar Chart

The final term frequencies were visualised using a horizontal bar chart to compare the relative prominence of the 20 most frequent terms.

```{r}
# Load packages for data visualisation
library(ggplot2)
library(RColorBrewer)
# Visualise the Top 20 most frequent terms
ggplot(
  rq1_top20_df,
  aes(
    x = reorder(display_term, frequency),
    y = frequency
  )
) +
  geom_col(fill = "steelblue") +
  geom_text(
    aes(label = frequency),
    hjust = -0.15,
    size = 3.5
  ) +
  coord_flip() +
  labs(
    title = "Top 20 Most Frequent Terms",
    subtitle = "Bluesky discussion of Tropical Cyclone Alfred",
    x = NULL,
    y = "Frequency",
    caption = "Related word forms were consolidated using stemming; readable labels are shown."
  ) +
  expand_limits(
    y = max(rq1_top20_df$frequency) * 1.12
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11),
    axis.text = element_text(size = 10),
    panel.grid.major.y = element_blank()
  )
```

The bar chart highlights the prominence of geographic and weather-related terms in the discussion. Terms associated with flooding, power, preparation, and information updates were also among the most frequent, showing that discussion extended beyond tracking the cyclone to include preparedness, impacts, and real-time information sharing.

### 5.4 - Word Cloud

A word cloud was created as a supplementary visualisation of the most frequent terms, with larger words representing higher term frequencies.

```{r}
# Load the package for word cloud visualisation
library(wordcloud)

# Create a readable word cloud from the final stemmed term frequencies
wordcloud(
  words = names(rq1_term_frequency),
  freq = rq1_term_frequency,
  min.freq = 200,
  max.words = 50,
  random.order = FALSE,
  colors = brewer.pal(8, "Dark2")
)
```
The word cloud provides a complementary visual summary of the dominant vocabulary. The prominence of geographic, weather, impact, and information-related terms is consistent with the patterns observed in the term-frequency analysis and bar chart.

## 6 - Contextual Interpretation

To support the interpretation of the term-frequency patterns without pre-defining themes, a reproducible random sample of 30 posts was selected from the full set of usable posts for contextual inspection.


```{r}
# Prepare all usable posts for contextual interpretation
rq1_context <- alfred_all %>%
  filter(text_usable) %>%
  select(author_handle, created_at, text)

# Check the number of posts available for contextual interpretation
dim(rq1_context)
```
The contextual dataset contained 6,792 posts with usable text, providing the full pool from which the random sample was drawn.

```{r}
# Set a seed so the random sample can be reproduced
set.seed(123)

# Randomly sample posts from the full usable-text dataset
rq1_context_sample <- rq1_context %>%
  slice_sample(n = 30)

# Check the sample size
dim(rq1_context_sample)
```
A fixed random seed was used to make the selection reproducible. The resulting sample contained 30 posts selected from the full usable-text dataset.

### 6.1 - Inspect Sampled Posts

The sampled posts were examined in their original text form to provide context for the patterns identified through term-frequency analysis.

```{r}
# Display the text of all 30 randomly sampled posts
options(tibble.width = Inf)

rq1_context_sample %>%
  arrange(created_at) %>%
  select(text) %>%
  print(n = 30, width = Inf)
```
The sampled posts contained recurring discussion of cyclone forecasts and changing landfall conditions, public warnings and preparation, flooding and other disruptions, and post-event recovery. The sample also included some posts that were primarily links or only indirectly related to the main discussion, which were not forced into the identified themes.

### 6.2 - Check Temporal Distribution

The distribution of the sampled posts across dates was examined to understand the temporal coverage of the random sample.

```{r}
# Check the temporal distribution of the random context sample
table(
  as.Date(rq1_context_sample$created_at)
)
```
The random sample was concentrated around the period of highest discussion activity, particularly from 4 to 9 March. This reflects the underlying dataset, where posting activity was substantially higher around the cyclone's approach, landfall, and immediate aftermath. The sample was therefore retained in its original random form rather than artificially balancing the number of posts across dates.

### 6.3 - Prepare Posts for Context Review

The sampled posts were organised chronologically and assigned post IDs to support systematic contextual review.

```{r}
# Prepare the random sample for systematic context review
rq1_context_review <- rq1_context_sample %>%
  arrange(created_at) %>%
  mutate(post_id = row_number()) %>%
  select(post_id, created_at, author_handle, text)

# Check the review dataset
dim(rq1_context_review)
```

```{r}
# Display reviewed posts with safe text encoding
rq1_context_review %>%
  mutate(text = iconv(text, from = "UTF-8",
                      to = "ASCII", sub = "")) %>%
  select(post_id, text) %>%
  print(n = 30, width = Inf)
```
The contextual review provided further support for 4 recurring information themes: **storm tracking and weather updates, preparedness and warnings, impacts and disruption, and post-event impacts and recovery**. Posts included changing forecasts and landfall conditions, preparation for possible power outages, flooding and transport disruption, and later discussion of clean-up and restoration. The sample was used to support interpretation of the broader term-frequency patterns rather than to estimate how common each theme was across the full dataset.

### RQ1 Summary

Overall, the combined term-frequency patterns and contextual inspection indicate four main information themes in Bluesky discussion of Tropical Cyclone Alfred: **storm tracking and weather updates, preparedness and warnings, impacts and disruption, and post-event impacts and recovery**. Discussion was geographically centred on Queensland, Brisbane and northern New South Wales, while the prominence of terms associated with news, live coverage and updates suggests that Bluesky also functioned as a channel for real-time information sharing during the event.