---
title: "Social Web Analytics of Tropical Cyclone Alfred on Bluesky"
author: "COMP3020 – Social Web Analytics"
date: "9 October 2026"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---
\begin{center}
\Large\textbf{Group 15}
\end{center}

\vspace{0.5cm}

\begin{center}
\renewcommand{\arraystretch}{1.6}
\begin{tabular}{|p{6cm}|p{3cm}|p{3cm}|}
\hline
\textbf{Group Members} & \textbf{Student ID} & \textbf{Contribution} \\
\hline
Phuong Linh Le & 22113467 & 34\% \\
\hline
Mai Khanh Ngoc Tran & 22180754 & 33\% \\
\hline
Tran Truc Thanh DINH & 22110241 & 33\% \\
\hline
\end{tabular}
\end{center}

\vspace{0.5cm}

**GitHub Repository:**

https://github.com/Stella264/COMP3020-Social-Web-Analytics-Group-Project

\newpage
The atrrr R package, an interface for the AT Protocol behind Bluesky, was utilised to collect Bluesky posts; the search term "Cyclone Alfred" was used for the collection and it spanned from February 23rd to March 16th, 2025. Data retrieval was conducted through daily search windows.
```{r, eval=FALSE}
# Load the package used for Bluesky data collection
library(atrrr)
```
```{r, eval=FALSE, include=FALSE}
# Initial API test - Retrieve a small sample to inspect the Bluesky data structure
alfred_test <- search_skeet(
  "Cyclone Alfred",
  sort = "latest",
  since = "2025-02-23",
  limit = 100
)
# Inspect the dimensions and available variables in the initial sample
dim(alfred_test)
names(alfred_test)
# Inspect key variables relevant to the project
head(alfred_test[, c(
  "created_at",
  "author_handle",
  "text",
  "reply_count",
  "repost_count",
  "like_count"
)])
# Check the actual temporal coverage returned by the API
range(alfred_test$created_at)
head(sort(alfred_test$created_at))
tail(sort(alfred_test$created_at))
```
```{r, eval=FALSE}
# Inspect the available arguments of search_skeet before defining the collection period
args(search_skeet)
```
```{r, eval=FALSE, include=FALSE}
# Test retrieval across the full event period using a single API request
alfred_event <- search_skeet(
  "Cyclone Alfred",
  sort = "latest",
  since = "2025-02-23",
  until = "2025-03-17",
  limit = 500
)

# Examine the coverage of the initial event-level retrieval
dim(alfred_event)
range(alfred_event$created_at, na.rm = TRUE)
length(unique(alfred_event$author_handle))
table(as.Date(alfred_event$created_at))
```
```{r, eval=FALSE, include=FALSE}
# 1. Check the dataset size
dim(alfred_event)
# 2. Check the actual date range
range(alfred_event$created_at, na.rm = TRUE)
# 3. Check the number of unique authors
length(unique(alfred_event$author_handle))
# 4. Check the number of posts per day
table(as.Date(alfred_event$created_at))
```
```{r, eval=FALSE, include=FALSE}
# Check the number of posts that are replies
sum(!is.na(alfred_event$in_reply_to))
# Calculate the total number of replies received
sum(alfred_event$reply_count, na.rm = TRUE)
# Calculate the total number of reposts
sum(alfred_event$repost_count, na.rm = TRUE)
# Calculate the total number of quotes
sum(alfred_event$quote_count, na.rm = TRUE)
# Check the number of posts containing mentions
sum(lengths(alfred_event$mentions) > 0, na.rm = TRUE)
# Check the number of posts containing links
sum(lengths(alfred_event$links) > 0, na.rm = TRUE)
```
```{r, eval=FALSE, include=FALSE}
# Test data retrieval for the earlier period of the event
alfred_before <- search_skeet(
  "Cyclone Alfred",
  sort = "latest",
  since = "2025-02-23",
  until = "2025-03-05",
  limit = 500
)
```
```{r, eval=FALSE, include=FALSE}
# Check the dataset size
dim(alfred_before)
# Check the actual date range returned
range(alfred_before$created_at, na.rm = TRUE)
# Check the number of unique authors
length(unique(alfred_before$author_handle))
```
```{r, eval=FALSE, include=FALSE}
# Test data retrieval using a one-day search window
alfred_23feb <- search_skeet(
  "Cyclone Alfred",
  sort = "latest",
  since = "2025-02-23",
  until = "2025-02-24",
  limit = 500
)
```
```{r, eval=FALSE, include=FALSE}
# Check the dataset size
dim(alfred_23feb)
# Check the actual date range returned
range(alfred_23feb$created_at, na.rm = TRUE)
# Check the number of unique authors
length(unique(alfred_23feb$author_handle))
```
```{r, eval=FALSE}
# Create a sequence of dates covering the full study period
dates <- seq(
  as.Date("2025-02-23"),
  as.Date("2025-03-16"),
  by = "day"
)
# Create an empty list to store the daily search results
alfred_daily <- vector("list", length(dates))
# Collect Bluesky posts separately for each day
for (i in seq_along(dates)) {
  # Define the start and end date for each daily search
  start_date <- dates[i]
  end_date <- start_date + 1
  # Display the date currently being collected
  cat("Collecting:", as.character(start_date), "\n")
  # Retrieve posts containing the search term "Cyclone Alfred"
  alfred_daily[[i]] <- search_skeet(
    "Cyclone Alfred",
    sort = "latest",
    since = as.character(start_date),
    until = as.character(end_date),
    limit = 1000
  )
}
```
```{r, eval=FALSE}
# Load dplyr for data manipulation
library(dplyr)
# Combine all daily search results into one final dataset
alfred_all <- bind_rows(alfred_daily)
# Check the size of the final dataset
dim(alfred_all)
# Check the actual date range of the final dataset
range(alfred_all$created_at, na.rm = TRUE)
# Check the number of unique authors
length(unique(alfred_all$author_handle))
# Check for duplicate posts based on URI
sum(duplicated(alfred_all$uri))
```
```{r, eval=FALSE}
# Check the number of posts collected per day
table(as.Date(alfred_all$created_at))
```
```{r, eval=FALSE}
# Check the number of posts that are replies
sum(!is.na(alfred_all$in_reply_to))
# Calculate the total number of replies received
sum(alfred_all$reply_count, na.rm = TRUE)
# Calculate the total number of reposts
sum(alfred_all$repost_count, na.rm = TRUE)
# Calculate the total number of quotes
sum(alfred_all$quote_count, na.rm = TRUE)
# Check the number of posts containing mentions
sum(lengths(alfred_all$mentions) > 0, na.rm = TRUE)
# Check the number of posts containing links
sum(lengths(alfred_all$links) > 0, na.rm = TRUE)
```
```{r, eval=FALSE}
# Save the final collected dataset for subsequent analysis
# file.exists("alfred_all.rds")
```
# Data Collection
# Data Understanding and Preparation
## Step 1 - Load Dataset
```{r, message=FALSE}
# Load required package
library(dplyr)
# Load the shared dataset
alfred_all_raw <- readRDS("Data/alfred_all.rds")
# Create a working copy
alfred_all <- alfred_all_raw
# Initial check
dim(alfred_all)
```
## Step 2 - Dataset Audit
```{r}
# Dataset class
# class(alfred_all)
# Dataset dimensions
dim(alfred_all)
# Variable names
# names(alfred_all)
# Variable types and structure
# dplyr::glimpse(alfred_all)
# Inspect first few records
# head(alfred_all)
# Check duplicate post identifiers
sum(duplicated(alfred_all$uri))
```
## Step 3 - Missingness and Record Suitability
```{r}
# Number of missing values in each variable
missing_values <- colSums(is.na(alfred_all))
# missing_values
sum(is.na(alfred_all$uri))
sum(is.na(alfred_all$author_handle))
sum(is.na(alfred_all$text))
sum(is.na(alfred_all$created_at))
sum(is.na(alfred_all$in_reply_to))
sum(is.na(alfred_all$quotes))
```
## Step 4 - User and Temporal Coverage
```{r}
# Number of unique authors
unique_users <- length(unique(alfred_all$author_handle))
unique_users
# Check created_at type
class(alfred_all$created_at)
# Actual post creation time range
date_range <- range(alfred_all$created_at,na.rm = TRUE)
date_range
# Number of posts by date
posts_per_day <- table(as.Date(alfred_all$created_at))
# posts_per_day
```
## Step 5 - Text Quality
```{r}
# Missing text
missing_text <- sum(is.na(alfred_all$text))
missing_text
# Blank / empty text
blank_text <- sum(trimws(alfred_all$text) == "",na.rm = TRUE)
blank_text
# Posts containing usable non-empty text
usable_text <- sum(
  !is.na(alfred_all$text) &
    trimws(alfred_all$text) != ""
)
usable_text
# Inspect language variable
# class(alfred_all$langs)
# str(alfred_all$langs)
# Number of language values attached to each post
# table(lengths(alfred_all$langs))
# Inspect actual language codes
language_values <- unlist(alfred_all$langs,use.names = FALSE)
# sort(
  # table(language_values),
  # decreasing = TRUE
# )
alfred_all$text_usable <-
  !is.na(alfred_all$text) &
  trimws(alfred_all$text) != ""
table(alfred_all$text_usable)
```
## Step 6 - Link Preparation
```{r}
# Inspect links variable
# class(alfred_all$links)
# str(alfred_all$links)
# Number of links recorded per post
# table(lengths(alfred_all$links))
# Number of posts with at least one recorded link
posts_with_links <- sum(lengths(alfred_all$links) > 0)
posts_with_links
link_rows <- which(lengths(alfred_all$links) > 0)
# head(alfred_all$links[link_rows],10)
alfred_all$has_link <-lengths(alfred_all$links) > 0
table(alfred_all$has_link)
prop.table(
  table(alfred_all$has_link)
)
```
## Step 7 - Network Feasibility
### 7A - Replies
```{r}
# Number of reply posts
reply_posts <- sum(
  !is.na(alfred_all$in_reply_to)
)
reply_posts
# Inspect reply identifiers
class(alfred_all$in_reply_to)
# head(na.omit(alfred_all$in_reply_to),10)
reply_rows <- which(!is.na(alfred_all$in_reply_to))
reply_parent_match <- match(
  alfred_all$in_reply_to[reply_rows],
  alfred_all$uri
)
reply_parents_in_dataset <- sum(!is.na(reply_parent_match))
reply_parents_in_dataset
reply_mapping_rate <-
  reply_parents_in_dataset /
  length(reply_rows)
reply_mapping_rate
reply_example <- data.frame(
  source_user =
    alfred_all$author_handle[reply_rows],
  target_uri =
    alfred_all$in_reply_to[reply_rows],
  target_user =
    alfred_all$author_handle[reply_parent_match]
)
# head(reply_example, 10)
```
### 7B - Mentions
```{r}
# Number of posts containing mentions
mention_posts <- sum(lengths(alfred_all$mentions) > 0)
mention_posts
# Inspect structure
# class(alfred_all$mentions)
# str(alfred_all$mentions)
# Inspect non-empty mention examples
mention_rows <- which(lengths(alfred_all$mentions) > 0)
# head(alfred_all$mentions[mention_rows],10)
# head(unlist(alfred_all$mentions[mention_rows],use.names = FALSE),20)
```
### 7C - Quotes
```{r}
# Inspect quotes variable
# class(alfred_all$quotes)
# str(alfred_all$quotes)
# Number of posts with quote information
quote_posts <- sum(!is.na(alfred_all$quotes))
quote_posts
# Inspect examples
# head(na.omit(alfred_all$quotes),10)
if (is.character(alfred_all$quotes)) {
  quote_rows <- which(
    !is.na(alfred_all$quotes)
  )
  quote_posts_in_dataset <- sum(
    alfred_all$quotes[quote_rows] %in%
      alfred_all$uri
  )
  quote_posts_in_dataset
}
```
## Step 8 - Final Preparation Summary
```{r}
prep_summary <- data.frame(
  Metric = c(
    "Total posts",
    "Total variables",
    "Unique users",
    "Duplicate URI",
    "Missing created_at",
    "Missing text",
    "Blank text",
    "Usable text posts",
    "Posts with links",
    "Reply posts",
    "Reply parents found in dataset",
    "Posts with mentions",
    "Posts with quote information"
  ),
  Value = c(
    nrow(alfred_all),
    ncol(alfred_all),
    unique_users,
    sum(duplicated(alfred_all$uri)),
    sum(is.na(alfred_all$created_at)),
    missing_text,
    blank_text,
    usable_text,
    posts_with_links,
    reply_posts,
    reply_parents_in_dataset,
    mention_posts,
    quote_posts
  )
)
prep_summary
date_range
```
# RQ1 - Content Themes

**Research Question:** What information themes characterised Bluesky discussion during Tropical Cyclone Alfred?

## 1 - Select Usable Text

For the content analysis, we only selected posts with valid textual content.
```{r}
# Select usable text for content analysis
rq1_text <- alfred_all$text[alfred_all$text_usable]
```
## 2 - Text Preprocessing

The chosen posts were transformed into a text corpus and processed prior to the term frequency analysis with the tm package. This package is an open-source tool for conducting text mining within R (Feinerer et al., 2008). All unsupported symbols, URLs, numbers, punctuation, common English stop words, and query-specific terms were removed from the data.
```{r, message=FALSE}
# Load the package for text processing
library(tm)
```
```{r, warning=FALSE}
# Create a text corpus
rq1_corpus <- Corpus(VectorSource(rq1_text))
# Convert text to ASCII to remove unsupported symbols and emojis
rq1_corpus <- tm_map(
  rq1_corpus,
  content_transformer(function(x) iconv(x, to = "ASCII", sub = ""))
)
```
```{r, warning=FALSE}
# 2.1 - Remove URLs from the corpus
rq1_corpus <- tm_map(
  rq1_corpus,
  content_transformer(function(x) gsub("http\\S+|www\\S+", "", x))
)
```
```{r, warning=FALSE}
# 2.2 - Convert all text to lowercase
rq1_corpus <- tm_map(rq1_corpus,content_transformer(tolower))
```
```{r, warning=FALSE}
# 2.3 - Remove numbers from the corpus
rq1_corpus <- tm_map(rq1_corpus,removeNumbers)
```
```{r, warning=FALSE}
# 2.4 - Remove punctuation from the corpus
rq1_corpus <- tm_map(rq1_corpus,removePunctuation)
```
```{r, warning=FALSE}
# 2.5 - Remove common English stopwords
rq1_corpus <- tm_map(rq1_corpus,removeWords,stopwords("english"))
```
```{r, warning=FALSE}
# 2.6 - Remove query-specific words
custom_stopwords <- c("cyclone", "alfred")
rq1_corpus <- tm_map(
  rq1_corpus,
  removeWords,
  custom_stopwords
)
```
```{r, warning=FALSE}
# 2.7 - Remove extra whitespace
rq1_corpus <- tm_map(rq1_corpus,stripWhitespace)
```
## 3 - Initial Term-Frequency Analysis

To identify the most frequent terms and decide whether additional text preprocessing, we built an initial TDM (Term Document Matrix), which supported by the text mining framework describe by Feinerer et al.(2008).
```{r}
# Create the initial Term-Document Matrix
rq1_tdm <- TermDocumentMatrix(rq1_corpus)
# Calculate initial term frequencies
rq1_term_frequency <- rowSums(as.matrix(rq1_tdm))
# Display the 20 most frequent terms
sort(rq1_term_frequency, decreasing = TRUE)[1:20]
```
The initial term frequency results helped detect limited analytical value words and to assess whether related word forms were being counted separately. So, these  checks informed the final text cleaning process before the main results were interpreted.

### 3.1 - Inspect Potentially Uninformative Terms

Frequently used words seemed to have less useful information were examined within the original posts before deciding if they should be removed.
```{r, echo=TRUE, results='hide'}
# Inspect the context of potentially uninformative frequent terms
terms_to_check <- c("will", "now", "make", "news", "live")
for (term in terms_to_check) {
  # Find posts containing the selected term
  matched_posts <- rq1_text[
    grepl(paste0("\\b", term, "\\b"), tolower(rq1_text))
  ]
  # Display the term and three example posts
  cat("\n\n--- TERM:", term, "---\n")
  print(head(matched_posts, 3))
}
```
This showed that *will*, *now*, and *make* were largely generic and provided limited information about the cyclone discussion. In contrast, *news* and *live* were retained as they frequently appeared in posts sharing news coverage and real time event updates. Therefore, *will*, *now*, and *make* were removed as additional stopwords.
```{r, warning=FALSE}
# Remove additional uninformative words identified through context inspection
additional_stopwords <- c("will", "now", "make")
rq1_corpus <- tm_map(rq1_corpus,removeWords,additional_stopwords)
# Remove extra whitespace created after removing additional words
rq1_corpus <- tm_map(rq1_corpus,stripWhitespace)
```
### 3.2 - Check Related Word Forms

To see whether different word forms representing the same concept were being counted separately, we examined different forms of frequent occurring words.
```{r}
# Check whether related word forms are splitting term frequencies
stem_check_terms <- c(
  "flood", "flooded", "flooding",
  "warn", "warned", "warning", "warnings",
  "prepare", "prepared", "preparing",
  "impact", "impacted", "impacts"
)
rq1_term_frequency[
  intersect(stem_check_terms, names(rq1_term_frequency))
]
```
From the result, we can see *flood*, *flooded*, and *flooding*, as well as different forms of *warning*, *prepare*, and *impact*, appeared as separate terms. Therefore, stemming was applied to consolidate related word forms and reduce fragmentation in the term frequency analysis, based on the approach described by Porter (1980).
```{r, warning=FALSE}
# Stem related word forms to a common root
rq1_corpus <- tm_map(rq1_corpus,stemDocument)
```
```{r, warning=FALSE}
# Remove residual query terms after stemming
rq1_corpus <- tm_map(rq1_corpus,removeWords,c("cyclonealfr", "tcalfr", "cyclon"))
rq1_corpus <- tm_map(rq1_corpus, stripWhitespace)

```
## 4 - Final Term-Frequency Analysis

Once the text preprocessing was completed the TDM and term frequencies were recalculated.
```{r}
# Recreate the Term-Document Matrix after final text preprocessing
rq1_tdm <- TermDocumentMatrix(rq1_corpus)
# Recalculate term frequencies after stemming
rq1_term_frequency <- rowSums(as.matrix(rq1_tdm))
# Display the 20 most frequent terms
# sort(rq1_term_frequency, decreasing = TRUE)[1:20]
```
The final frequency results show that Bluesky discussion was mainly focused on the cyclone's geographic location, development, and impacts. Geographic terms such as: *Queensland*, *Brisbane*, *coast*, and *NSW*. They suggest substantial attention to areas affected or threatened by the cyclone. Weather and eventrelated terms such as *tropical*, *storm*, *wind*, *landfall*, and *rain* were also prominent, reflecting ongoing discussion of the cyclone's movement and conditions. Terms including *flood*, *power*, and *prepare* suggest attention to potential or actual consequences and preparedness, while *news*, *live*, and *update* suggest that real time information sharing was also prominent.

## 5 - Visualisation

### 5.1 - Prepare Top 20 Terms

```{r}
# Prepare the Top 20 frequent terms for visualisation
rq1_top20 <- sort(rq1_term_frequency,decreasing = TRUE)[1:20]
# Convert the Top 20 terms into a data frame
rq1_top20_df <- data.frame(term = names(rq1_top20),
  frequency = as.numeric(rq1_top20)
)
# Display the prepared data
# rq1_top20_df
```
### 5.2 - Create Readable Labels

Because stemming produces shortened word forms, readable labels were used in the visualisations without changing the original calculated frequencies.
```{r}
# Create readable labels for stemmed terms in visualisations
rq1_top20_df$display_term <- c(
  "Tropical",
  "Queensland",
  "Coast",
  "Brisbane",
  "Australia",
  "News",
  "Flood",
  "March",
  "NSW",
  "Live",
  "Wind",
  "Storm",
  "Landfall",
  "Ex-tropical",
  "Rain",
  "Power",
  "Update",
  "Hit",
  "South",
  "Prepare"
)

# Check that display labels match the original stemmed terms
# rq1_top20_df[, c("term", "display_term", "frequency")]
```
### 5.3 - Top 20 Term-Frequency Bar Chart

A horizontal bar chart was used to compare the relative prominence of the 20 most frequent terms.
```{r, message=FALSE}
# Load packages for data visualisation
library(ggplot2)
library(RColorBrewer)
# Visualise the Top 20 most frequent terms
ggplot(
  rq1_top20_df,
  aes(
    x = reorder(display_term, frequency),
    y = frequency
  )
) +
  geom_col(fill = "steelblue") +
  geom_text(
    aes(label = frequency),
    hjust = -0.15,
    size = 3.5
  ) +
  coord_flip() +
  labs(
    title = "Top 20 Most Frequent Terms",
    subtitle = "Bluesky discussion of Tropical Cyclone Alfred",
    x = NULL,
    y = "Frequency",
    caption = "Related word forms were consolidated using stemming; readable labels are shown."
  ) +
  expand_limits(
    y = max(rq1_top20_df$frequency) * 1.12
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11),
    axis.text = element_text(size = 10),
    panel.grid.major.y = element_blank()
  )
```
The bar chart highlights the high frequency of geographic and weather-related terms in the discussion. Terms associated with flooding, power, preparation, and information updates also appeared frequently, suggesting that discussions were not limited to tracking the cyclone but also covered preparation, impacts, and real-time information sharing.

### 5.4 - Word Cloud

A word cloud was created as a supplementary visualisation of the most frequent words, where larger text representing higher term frequencies.
```{r}
library(wordcloud)
# Create word cloud data from RQ1 term frequencies
rq1_wc <- data.frame(
  term = names(rq1_term_frequency),
  frequency = as.numeric(rq1_term_frequency)
)
# Apply readable labels from Top 20
rq1_wc$display_term <- rq1_wc$term
matched <- match(rq1_wc$term, rq1_top20_df$term)
rq1_wc$display_term[!is.na(matched)] <-
  rq1_top20_df$display_term[matched[!is.na(matched)]]
# Additional readable labels
extra_labels <- c(
  "prepar" = "Preparation",
  "resid" = "Residents",
  "peopl" = "People",
  "extrop" = "Ex-tropical",
  "updat" = "Update",
  "forecast" = "Forecast",
  "warn" = "Warning",
  "southeast" = "Southeast"
)
idx <- match(rq1_wc$term, names(extra_labels))
rq1_wc$display_term[!is.na(idx)] <-unname(extra_labels[idx[!is.na(idx)]])
# Select top 40 terms
rq1_wc <- rq1_wc[rq1_wc$term != "cyclonealfr",]
rq1_wc <- rq1_wc[order(-rq1_wc$frequency), ]
rq1_wc <- head(rq1_wc, 40)
# Colour palette
wc_palette <- c(
  "#F29D38",
  "#E96C91",
  "#16A6A0",
  "#8054B8",
  "#238BC4",
  "#183F78"
)
set.seed(123)
word_colors <- sample(
  rep(wc_palette, length.out = nrow(rq1_wc))
)
# Create word cloud
set.seed(123)
wordcloud(
  words = rq1_wc$display_term,
  freq = rq1_wc$frequency,
  scale = c(3.0, 0.85),
  min.freq = 0,
  max.words = 40,
  random.order = FALSE,
  rot.per = 0,
  colors = word_colors,
  ordered.colors = TRUE,
  family = "sans"
)
```
The word cloud provides a complementary visual summary of the dominant vocabulary. The frequent appearance of terms related to locations, weather, impacts, and information is consistent with the patterns observed in the term-frequency analysis and bar chart.

## 6 - Contextual Interpretation

To support the interpretation of the term-frequency patterns without pre-defining themes,a reproducible random sample of 30 posts from the full set of usable posts was selected for reviewing their original context.
```{r}
# Prepare all usable posts for contextual interpretation
rq1_context <- alfred_all %>%
  filter(text_usable) %>%
  select(author_handle, created_at, text)
# Check the number of posts available for contextual interpretation
dim(rq1_context)
```
Then, 6,792 posts with usable text, providing the full pool used for random sampling.
```{r}
# Set a seed so the random sample can be reproduced
set.seed(123)
# Randomly sample posts from the full usable-text dataset
rq1_context_sample <- rq1_context %>%slice_sample(n = 30)
# Check the sample size
dim(rq1_context_sample)
```
A fixed random seed was set to make the selection reproducible. The resulting sample consisted of 30 posts drawn from the full usable-text dataset.

### 6.1 - Inspect Sampled Posts

The sampled posts were reviewed using their original textual content to provide context for the trends observed in the word-frequency analysis.
```{r, echo=TRUE, results='hide'}
# Display the text of all 30 randomly sampled posts
options(tibble.width = Inf)
rq1_context_sample %>%
  arrange(created_at) %>%
  select(text) %>%
  print(n = 30, width = Inf)
```
The review of the sampled posts revealed several recurring topics, including: cyclone forecasts and changing landfall conditions; public warnings and preparation; flooding and other disruptions, and post-event recovery. The sample also included some posts that were primarily links or only indirectly related to the main discussion, which were not forced into the identified themes.

### 6.2 - Check Temporal Distribution

The posting dates of the 30 sampled posts were examined to understand the temporal coverage of the random sample.
```{r}
# Check the temporal distribution of the random context sample
table(as.Date(rq1_context_sample$created_at))
```
Most of the sampled posts were published between 4 and 9 March. This reflects the original dataset, where posting activity was substantially higher around the cyclone's approach, landfall, and immediate aftermath. The sample was therefore retained in its original random form rather than artificially balancing the number of posts across dates.

### 6.3 - Prepare Posts for Context Review

To facilitate a structured review, the sampled posts were organised chronologically and assigned post IDs.
```{r, echo=TRUE, results='hide'}
# Prepare the random sample for systematic context review
rq1_context_review <- rq1_context_sample %>%
  arrange(created_at) %>%
  mutate(post_id = row_number()) %>%
  select(post_id, created_at, author_handle, text)
# Check the review dataset
dim(rq1_context_review)
```
```{r, echo=TRUE, results='hide'}
# Display reviewed posts with safe text encoding
rq1_context_review %>%
  mutate(text = iconv(text, from = "UTF-8",
                      to = "ASCII", sub = "")) %>%
  select(post_id, text) %>%
  print(n = 30, width = Inf)
```
The review of original posts offered additional evidence for four recurring information themes: **storm tracking and weather updates, preparedness and warnings, impacts and disruption, and post-event impacts and recovery**. The content covered changing forecasts and landfall conditions, preparation for possible power outages, flooding and transport disruption, and later discussion of cleaning and recovery activities. The sample was used to assist in explaining of the broader term-frequency patterns instead of measuring how common each theme was across the full dataset.

### RQ1 Summary

Taken together, the combined term-frequency patterns and contextual inspection indicate four key information themes in Bluesky discussion of Tropical Cyclone Alfred: **storm tracking and weather updates, preparedness and warnings, impacts and disruption, and post-event impacts and recovery**. Discussion was geographically centred on Queensland, Brisbane and northern New South Wales, while the prominence of terms associated with news, live coverage and updates suggests that Bluesky also functioned as a channel for real-time information sharing throughout the cyclone event.

### RQ1 Limitations

Several limitations should be considered in this analysis. First, term-frequency analysis identifies frequently used words but cannot completely reflect their meanings within the original posts. According to Grimmer and Stewart (2013), automated text analysis requires careful interpretation and validation rather than relying solely on numerical results. Although stemming helps consolidate related word forms, it may also reduce linguistic detail.Second, the contextual interpretation was based on a random sample of 30 posts from 6,792 usable posts. Therefore, the four identified themes should be understood as repeated patterns supported by the analysis, rather than exhaustive categories or estimates of theme prevalence. Finally, the dataset was limited to Bluesky posts collected using cyclone-related search terms during the selected study period, so the findings may not fully reflect all public conversations about Tropical Cyclone Alfred.